A demonstration of the \LaTeXe\ class file for \itshape{SAGE Publications}

### 2015-12-22 body/appendix_algo.tex
andrea's modifications

First, we define an abtract structure State,
that describes a contact configuration.
The use of queues allows a FIFO approach regarding the order 
in which contacts are tested: we try to replace older contacts first when necessary.
Thus the algorithm is deterministic even though it can handle acyclic motions.

@@

>>>>

First, we define an abstract structure State,
that describes a contact configuration.
The use of queues allows a FIFO approach regarding the order 
in which contacts are tested: we try to replace older contacts first when necessary.
Thus the algorithm is deterministic even though it can handle acyclic motions.

@@

\section{Derivation of the manipulability ellipsoid}
\label{app:manipulability}

Again, we assume that $\mathbf{J}$ is full rank. We discard the $k$ indices, and write the pseudo-inverse of $\mathbf{J}$ as $\mathbf{J}^{\dagger}$.
\begin{eqnarray*}
\mathbf{\dot{p}} & =  & \mathbf{J} \dot{\mathbf{q}} \\ 
\mathbf{\dot{q}} & =  & \mathbf{J}^{\dagger} \dot{\mathbf{p}} \\ 
\mathbf{\dot{q}}^T & =  & \dot{\mathbf{p}}^T \mathbf{J}^{\dagger T} \\ 
\mathbf{\dot{q}}^T\mathbf{\dot{q}} & =  & \dot{\mathbf{p}}^T \mathbf{J}^{\dagger T} \mathbf{J}^{\dagger} \dot{\mathbf{p}}\\ 
\end{eqnarray*}

Then, proving that $\mathbf{J}^{\dagger T} \mathbf{J}^{\dagger} = (\mathbf{J}\mathbf{J}^T)^{-1}$ is straight-forward,
using the SVD decomposition of each term.

### 2015-12-22 body/conclusion.tex
andrea's modifications

In this paper we consider acyclic contact planning in cluttered environments, formulated as two sub problems that we address sequentially.
The first problem is \Pa, that consists in computing a guide path for the root of the robot\deleted{ that can be extended}; the second problem is \Pb, that is the computation of a discrete sequence of contacts along the root path.
Our contribution to \Pa is a generic characterization of the properties that the guide path must satisfy, in particular to enforce the completeness of the acyclic contact planner. We introduced a low-dimensional space $C_{reach}$ that can be mapped 
into the contact submanifold of the robot, approximated, and efficiently sampled by our Reachability-Based planner.
Our contribution to \Pb is a pragmatic contact generation scheme that can take into
account criteria to enforce interesting properties on the generated contacts (such as robustness, energy efficiency or naturalness).

Aside from the theoretical contributions, our results demonstrate that our method allows a very interesting compromise between three 
criteria that are hard to conciliate: generality, performance, and quality of the solution, making it the first acyclic contact
planner compatible with interactive applications.
%
\textbf{Regarding generality}, the reachability condition, coupled with an approach based on limb decomposition, 
allows the method to address arbitrary multiped robots. The only pre-requisite is the specification 
of the volumes $W^0$.
%
\textbf{Regarding performance}, our framework outperforms existing methods in addressing either \Pa or \Pb, \adnote{saying you outperform existing methods when you do not show any comparison is risky, the reviewers might ask for comparisons} leading to computation costs close to real-time in
\replaced{known static}{statically known} environments.
%
\textbf{Regarding the quality of the paths}, a parametrization of the reachability condition allows us to compute
relevant paths in all the presented scenarios, with low rejection rates.
As for \cite{Bouyarmane2009}, failures can still occur, due to the compromise criterion used in computing the guide path.
The low computational burden of our framework however allows for fast replanning \replaced{in case of failure}{when the solution fails}.

One direction for future work is to focus on a more accurate formulation of $C_{reach}$ to address 
this limitation.
Our main objective is of course to generate the complete interpolation between the contact sequences.
We have already been able to achieve this for HRP-2 for some of the computed sequences~\citep{Carpentier2016}, but we \replaced{admit}{concede} that additional
work is required on the planning side to obtain a seamless workflow. To achieve this, we are currently working on the notion of transition certificate, i.e. formulating
conditions that guarantee that the interpolation between two contact configurations is dynamically feasible. We are also
working on new heuristics more closely related to the dynamics of the systems (for instance, choose the contacts that minimize the jerk between two contact configurations).
\adnote{The 'future work' part seems a bit poor to me}

>>>>

In this paper we consider acyclic contact planning in cluttered environments, formulated as two sub problems that we address sequentially.
The first problem is \Pa, that consists in computing a guide path for the root of the robot that can be extended; the second problem is \Pb, that is the computation of a discrete sequence of contacts along the root path.
Our contribution to \Pa is a generic characterization of the properties that the guide path must satisfy, in particular to enforce the completeness of the acyclic contact planner. We introduced a low-dimensional space $C_{reach}$ that can be mapped 
into the contact submanifold of the robot, approximated, and efficiently sampled by our Reachability-Based planner.
Our contribution to \Pb is a pragmatic contact generation scheme that can take into
account criteria to enforce interesting properties on the generated contacts (such as robustness, energy efficiency or naturalness).

Aside from the theoretical contributions, our results demonstrate that our method allows a very interesting compromise between three 
criteria that are hard to conciliate: generality, performance, and quality of the solution, making it the first acyclic contact
planner compatible with interactive applications.
%
\textbf{Regarding generality}, the reachability condition, coupled with an approach based on limb decomposition, 
allows the method to address arbitrary multiped robots. The only pre-requisite is the specification 
of the volumes $W^0$.
%
\textbf{Regarding performance}, our framework is really efficient in addressing both \Pa and \Pb. This results in computation costs close to real-time in
known static environments.
%
\textbf{Regarding the quality of the paths}, a parametrization of the reachability condition allows us to compute
relevant paths in all the presented scenarios, with low rejection rates.
As for \cite{Bouyarmane2009}, failures can still occur, due to the compromise criterion used in computing the guide path.
The low computational burden of our framework however allows for fast replanning in case of failure.

One direction for future work is to focus on a more accurate formulation of $C_{reach}$ to address 
this limitation.
Our main objective is of course to generate the complete interpolation between the contact sequences.
We have already been able to achieve this for HRP-2 for some of the computed sequences~\citep{Carpentier2016}, but we admit that additional
work is required on the planning side to obtain a seamless workflow. To achieve this, we are currently working on the notion of transition certificate, i.e. formulating
conditions that guarantee that the interpolation between two contact configurations is dynamically feasible. We are also
working on new heuristics more closely related to the dynamics of the systems (for instance, choose the contacts that minimize the jerk between two contact configurations).
\adnote{The 'future work' part seems a bit poor to me}

### 2015-12-22 body/contact.tex
andrea's modifications

% !TEX root =  ../main.tex
\section{From a guide path to a discrete sequence of contact configurations ($\mathcal{P}_2$)}
\label{sec:contact}
As an input of this stage, we consider a truly feasible guide path $\mathbf{q}^0(t) : [0,1] \longrightarrow SE(3)$ for the root of the robot $R$. We now consider the second problem of computing a discrete sequence of contact configurations $\mathbf{Q}^{\overline{0}}$ for the limbs of the robot along $\mathbf{q}^0(t)$. \deleted[remark={the input is not truly feasible, actually it is not even geometrically feasible, it is only approximately geometrically feasible}]{From true feasibility we know that such a trajectory exists.}

In this Section we first describe a single contact-generation process, that is how to generate a contact configuration for a given limb 
and root location.
Then, on this method we build an iterative algorithm to generate a discrete sequence of contact configurations in equilibrium\deleted{, along a truly feasible path}.

@@

\subsection{Definition of a contact generator}
\label{sec:single_contact}
We consider a limb $R^k$  and a root configuration $\mathbf{q}^0$\deleted{ truly feasible regarding $R^k$}. FIGURE
From true feasibility we know that there exists an infinite list of mappings $P = [\pi_0,..,\pi_i, ...]$ that verify \eqref{eq:pi} for $\mathbf{q}^0$.
The issue is then to evaluate locally at least one element of $P$ at $\mathbf{q}^0$.

While exhibiting analytically a $\pi_i$ does not seem tractable, we can iteratively evaluate $\pi_i(\mathbf{q}^0)$ as follows:
\begin{enumerate}
\item Generate randomly a \replaced{collision-free}{valid} limb configuration;
\item Project the end-effector onto the closest surface with inverse kinematics;
\item If a valid solution is found, stop. Otherwise repeat from step 1.
\end{enumerate}

@@

We first define $C_{Contact}^{\epsilon} \supset C_{Contact}$ as the set of configurations almost in contact. This means that the minimal distance 
between an effector and a surface of the environment is less than a small value $\epsilon$.
We then apply the following steps:
\begin{enumerate}
\item Generate offline $N$ valid sample limb configurations $\mathbf{q}^k_i,  0 \leq i < N$;
\item Using the end-effector positions $\mathbf{p}(\mathbf{q}^k_i)$ as indices, store each sample in an octree stata structure;
\item Figure~\ref{fig:contact_gen}--2 and 3--:\adnote{This notation for subfigures is not clear, consider using subfigures, e.g. 2a, 2b} At runtime, when contact creation is required, retrieve from the octree the list of samples $S \subset C_{Contact}^{\epsilon}$ close to contact;
\item According to a user-defined heuristic $h$, sort the elements of $S$;
\item Figure~\ref{fig:contact_gen}--4 and 5--: Select the first configuration of $S$ (or fail if $S$ is empty). Project the configuration onto contact with inverse kinematics;
\item If a valid \adnote{being valid depends also on the other contacts because it regards also equilibrium, doesn't it? Then this assessment depends on the other contacts previously generated. Doesn't this have implications on the completeness? Even if it doesn't, isn't it worth to talk about it?} solution is found, stop. Otherwise remove the element from $S$ and go back step 5.
\end{enumerate}

This approach has two main advantages.
First, this allows us to select a large number of candidates in a single proximity request.
Having several candidates is interesting, because it allows to compare them \replaced{using}{regarding} a user-selected heuristic $h$, thus obtaining
a locally-optimal candidate.
Furthermore, the fact that the candidates are already close to contact increases the odds that the inverse kinematics will converge to a valid solution in a small number of iterations.
Regarding completeness, it is immediate to verify that as $N$ and $\epsilon$ grow, the completeness of the method is restored.
$N$ is a parameter that allows us to specify the trade-off between completeness and efficiency.
The reader is invited to refer to our previous work for an extensive discussion on the optimal value of $N$ \citep{Tonneau2014}.

We thus have defined a real-time\deleted{, extremely efficient} contact generator that can be as complete as required by the context.
%We compute the image of a mapping $f$ along $\mathbf{q}^0(t)$ into a discrete sequence of contacts $ \in C_{contact}$. We first describe the discretization of $\mathbf{q}^0(t)$ and the algorithm used to create or break contacts along it. Then we describe how contacts are explicitely generated along the discretized trajectory.

@@

As an input, we consider the truly feasible guide path $\mathbf{q}^0(t)$, automatically discretized into a sequence of $j$ key placements:  
\begin{equation*}
	\mathbf{Q}^0 = [\mathbf{q}^0_{0}; \mathbf{q}^0_{i}; ..., \mathbf{q}^0_{j-1}]
\end{equation*} 
where $\mathbf{q}^0_{0}$ and $\mathbf{q}^0_{j-1}$ respectively correspond to the start and goal configurations. To ensure continuity in the contact transition phases, we rewrite $\pi$ under the following recursive form for any $0<i<j$:
\begin{equation*}
    \pi\colon\left\{
    \begin{aligned}		
        \mathbf{Q}^0 \in C_{reach} & \longrightarrow C_{contact} \\
        %\mathbf{q}^{0}_0 &  \longrightarrow  \mathbf{q}_{start} \\
        \mathbf{q}^{0}_i &  \longrightarrow  \mathbf{q}^{0}_i \oplus g(\mathbf{q}_{i - 1},\mathbf{q}^{0}_i) 
    \end{aligned}
    \right.
\end{equation*} 
\adnote{Not clear, what is g?}
We initialize the recurrence with $\pi(\mathbf{q}^0_{0}) = \mathbf{q}_0$ the initial configuration of the robot.

At each step, we consider a new root placement $\mathbf{q}^0_{i}$ and try to generate a contact configuration that preserves as much as possible
the previous contacts while allowing for static equilibrium.

@@

Fig.~\ref{fig:break_contact} illustrates the contact-persistence strategy.
If possible, a limb in contact at time $i-1$ remains in contact at time $i$. The contact is broken if an inverse-kinematics solver fails to find a collision-free limb configuration that satisfies joint limits. The solver is directly provided by the HPP software.

@@

%\subsection{Generation of a contact configuration}  
\subsection{Creating contacts.}
Contacts are created using three rules:
\begin{enumerate}
\item Only one contact creation can happen between two consecutive states. This increases the odds that the interpolation between the two states \deleted[remark={not 100\% sure}]{will} be feasible; \adnote{Before you said that you ensured that transition between two states is feasible...}
\item A contact is validated if and only if the resulting configuration is in static equilibrium, no matter what user-defined heuristics are used; 
\item We use a FIFO approach:  we always try first to create a contact with the limb that has been contact-free the longest. If the contact creation
was not successful for a limb, the limb is pushed on top of the queue, and will only be tried again after the others have been treated.
\end{enumerate}

These three rules provide a deterministic manner to generate contacts along the discretized guide path.
FIGURE. For a given truly feasible path, our current implementation does not guarantee that the planner will succeed in generating a configuration in static equilibrium, because
of the deterministic approach used to break the combinatorial.
However, in practice the planner is successful in the large majority of cases, as discussed in Section~\ref{sec:stat1}, thanks to 
the heuristics that we use for contact selection.

>>>>

% !TEX root =  ../main.tex
\section{From a guide path to a discrete sequence of contact configurations ($\mathcal{P}_2$)}
\label{sec:contact}
As an input of this stage, we consider a geometrically feasible guide path $\mathbf{q}^0(t) : [0,1] \longrightarrow SE(3)$ for the root of the robot $R$. We now consider the second problem of computing a discrete sequence of contact configurations $\mathbf{Q}^{\overline{0}}$ for the limbs of the robot along $\mathbf{q}^0(t)$.

In this Section we first describe a single contact-generation process, that is how to generate a contact configuration for a given limb 
and root location.
Then, on this method we build an iterative algorithm to generate a discrete sequence of contact configurations in equilibrium.

@@

\subsection{Definition of a contact generator}
\label{sec:single_contact}
We consider a limb $R^k$  and a root configuration $\mathbf{q}^0$. TODO FIGURE
From true feasibility we know that there exists an infinite list of mappings $P = [\pi_0,..,\pi_i, ...]$ that verify \eqref{eq:pi} for $\mathbf{q}^0$.
The issue is then to evaluate locally at least one element of $P$ at $\mathbf{q}^0$.

While exhibiting analytically a $\pi_i$ does not seem tractable, we can iteratively evaluate $\pi_i(\mathbf{q}^0)$ as follows:
\begin{enumerate}
\item Generate randomly a collision-free limb configuration;
\item Project the end-effector onto the closest surface with inverse kinematics;
\item If a valid solution is found, stop. Otherwise repeat from step 1.
\end{enumerate}

@@

We first define $C_{Contact}^{\epsilon} \supset C_{Contact}$ as the set of configurations almost in contact. This means that the minimal distance 
between an effector and a surface of the environment is less than a small value $\epsilon$.
We then apply the following steps:
\begin{enumerate}
\item Generate offline $N$ valid sample limb configurations $\mathbf{q}^k_i,  0 \leq i < N$;
\item Using the end-effector positions $\mathbf{p}(\mathbf{q}^k_i)$ as indices, store each sample in an octree stata structure;
\item Figure~\ref{fig:contact_gen}--b and c--: At runtime, when contact creation is required, retrieve from the octree the list of samples $S \subset C_{Contact}^{\epsilon}$ close to contact;
\item According to a user-defined heuristic $h$, sort the elements of $S$;
\item Figure~\ref{fig:contact_gen}--d and e--: Select the first configuration of $S$ (or fail if $S$ is empty). Project the configuration onto contact with inverse kinematics;
\item If a valid solution is found, stop. Otherwise remove the element from $S$ and go back step 5.
\end{enumerate}

This approach has two main advantages.
First, this allows us to select a large number of candidates in a single proximity request.
Having several candidates is interesting, because it allows to compare them using a user-selected heuristic $h$, thus obtaining
a locally-optimal candidate.
Furthermore, the fact that the candidates are already close to contact increases the odds that the inverse kinematics will converge to a valid solution in a small number of iterations.
Regarding completeness, it is immediate to verify that as $N$ and $\epsilon$ grow, the completeness of the method is restored.
$N$ is a parameter that allows us to specify the trade-off between completeness and efficiency.
The reader is invited to refer to our previous work for an extensive discussion on the optimal value of $N$ \citep{Tonneau2014}.

We thus have defined a real-time contact generator. For a given configuration, the generator can be as complete as required by the context (by increasing the number of samples).

@@

As an input, we consider the guide path $\mathbf{q}^0(t)$, automatically discretized into a sequence of $j$ key placements:  
\begin{equation*}
	\mathbf{Q}^0 = [\mathbf{q}^0_{0}; \mathbf{q}^0_{i}; ..., \mathbf{q}^0_{j-1}]
\end{equation*} 
where $\mathbf{q}^0_{0}$ and $\mathbf{q}^0_{j-1}$ respectively correspond to the start and goal configurations. To ensure continuity in the contact transition phases, we rewrite $\pi$ under the following recursive form for any $0<i<j$:
\begin{equation*}
    \pi\colon\left\{
    \begin{aligned}		
        \mathbf{Q}^0 \in C_{reach} & \longrightarrow C_{contact} \\
        %\mathbf{q}^{0}_0 &  \longrightarrow  \mathbf{q}_{start} \\
        \mathbf{q}^{0}_i &  \longrightarrow  \mathbf{q}^{0}_i \oplus g(\mathbf{q}_{i - 1},\mathbf{q}^{0}_i) 
    \end{aligned}
    \right.
\end{equation*} 
$g$ is the method that extends a root configuration into a full-body configuration. At each step, it tries to generate a contact configuration that preserves as much as possible
the previous contacts while allowing for static equilibrium. The objective is thus to characterize $g$.
We initialize the recurrence with $\pi(\mathbf{q}^0_{0}) = \mathbf{q}_0$ the initial configuration of the robot.

@@

Figure~\ref{fig:break_contact} illustrates the contact-persistence strategy.
If possible, a limb in contact at time $i-1$ remains in contact at time $i$. The contact is broken if an inverse-kinematics solver fails to find a collision-free limb configuration that satisfies joint limits. The solver is directly provided by the HPP software.

@@

%\subsection{Generation of a contact configuration}  
\subsection{Creating contacts.}
Contacts are created using three rules:
\begin{enumerate}
\item Only one contact creation can happen between two consecutive states. This increases the odds that the interpolation between the two states be feasible;
%\adnote{Before you said that you ensured that transition between two states is feasible...}
\item A contact is validated if and only if the resulting configuration is in static equilibrium, no matter what user-defined heuristics are used; 
\item We use a FIFO approach:  we always try first to create a contact with the limb that has been contact-free the longest. If the contact creation
was not successful for a limb, the limb is pushed on top of the queue, and will only be tried again after the others have been treated.
\end{enumerate}

These three rules provide a deterministic manner to generate contacts along the discretized guide path.
TODO FIGURE. For a given truly feasible path, our current implementation does not guarantee that the planner will succeed in generating a configuration in static equilibrium, because
of the deterministic approach used to break the combinatorial.
However, in practice the planner is successful in the large majority of cases, as discussed in Section~\ref{sec:perf}, thanks to 
the heuristics that we use for contact selection.

### 2015-12-22 body/efort.tex
andrea's modifications

From \eqref{jac} we can obtain the following equality: \adnote{how?}
\begin{equation} \label{jaceq}
\mathbf{\dot{p}}^T(\mathbf{J}\mathbf{J}^T)^{-1}\mathbf{\dot{p}} = \dot{\mathbf{q}}^T \dot{\mathbf{q}}
\end{equation}

@@

\begin{equation} \label{ellipsoid}
\mathbf{\dot{p}}^T(\mathbf{J}\mathbf{J}^T)^{-1}\mathbf{\dot{p}} \leq 1
\end{equation}
This ellipsoid is called the manipulability ellipsoid, or velocity ellipsoid, introduced by \cite{Yoshikawa1984}. It describes the set of velocities that can
be reached under the constraint \eqref{ball} for the current configuration.
The longer an axis of the ellipsoid is, the faster the configuration can move along the direction of the axis.
Fig.~\ref{sec:efort_ellipsoid} - left shows the velocity ellipsoid for different configurations of a manipulator with two degrees of freedom.

@@

Exploiting \eqref{jac}, we can easily see that the set of achievable forces in $\mathbb{R}^m$ subject to the constraint:
\begin{equation*} \label{ballforce}
||\mathbf{\tau}||^2 \leq 1
\end{equation*}
is the so-called force ellipsoid (Fig.~\ref{sec:efort_ellipsoid} - right), defined by:
\begin{equation} \label{ellipsoidforce}
\mathbf{f}^T (\mathbf{J}\mathbf{J}^T) \mathbf{f} \leq 1
\end{equation}

\subsubsection{Manipulability-based heuristics.}
From these definitions, we can derive three useful heuristics, that all account \replaced{for}{from} the environment and the task being performed.
The first one, EFORT, was introduced by \cite{Tonneau2014}; the other two are new minor contributions, derived from these previous works.

@@

$h_{EFORT}$ and $h_{vel}$ will favor contacts that allow large efforts or fast modifications in the velocity.
$EFORT$ in particular is useful for tasks such as standing up, pushing / pulling.
In other less demanding cases, manipulability can also be considered \deleted{regarding its original purpose, that means} to avoid singularities.
To do so, we can consider the manipulability measure $w$, also given by \citeauthor{Yoshikawa1984}:

\begin{equation*} \label{ellipsoid}
w(\mathbf{q}) = \sqrt{\mathbf{J}\mathbf{J}^T}
\end{equation*} \adnote{Didn't you forget something in this equation? I guess $w$ is a scalar, but the right side is a matrix.}
$w$ measures the ``distance'' between a given configuration and a singular configuration. When $w$ is equal to 0, the configuration is in a singular state;
the greater $w$ is, the further away the configuration is from singularity.
We then derive:
\begin{equation}
h_{w}(\mathbf{q}, \mathbf{m}) = \sqrt{\mathbf{J}\mathbf{J}^T} ( \mu_0 \mathbf{n}^T \mathbf{m})
\end{equation}

$h_{EFORT}$, $h_{vel}$ and $h_{w}$ define three kinematic heuristics, fast to compute (indeed, thanks to our sampling-based approach, the Jacobian and inverse products
can be precomputed offline), that allow the planner to select the \replaced{best}{more relevant}
candidates \replaced{with respect to}{regarding} the local motion task along the guide path.
This results in faster computation times, and a reduced number of overall contact switches.

>>>>

We assume that $J$ is full rank (this is reasonable since singular configurations are discarded by our algorithm).
From \eqref{jac} we can thus obtain the following equality (Appendix~\ref{app:manipulability}):
\begin{equation} \label{jaceq}
\mathbf{\dot{p}}^T(\mathbf{J}\mathbf{J}^T)^{-1}\mathbf{\dot{p}} = \dot{\mathbf{q}}^T \dot{\mathbf{q}}
\end{equation}

@@

\begin{equation} \label{ellipsoid}
\mathbf{\dot{p}}^T(\mathbf{J}\mathbf{J}^T)^{-1}\mathbf{\dot{p}} \leq 1
\end{equation}
This ellipsoid is called the manipulability ellipsoid, or velocity ellipsoid, introduced by \cite{Yoshikawa1984}. It describes the set of velocities that can
be reached under the constraint \eqref{ball} for the current configuration.
The longer an axis of the ellipsoid is, the faster the configuration can move along the direction of the axis.
Figure~\ref{sec:efort_ellipsoid} - left shows the velocity ellipsoid for different configurations of a manipulator with two degrees of freedom.

@@

Exploiting \eqref{jac}, we can easily see that the set of achievable forces in $\mathbb{R}^m$ subject to the constraint:
\begin{equation*} \label{ballforce}
||\mathbf{\tau}||^2 \leq 1
\end{equation*}
is the so-called force ellipsoid (Figure~\ref{sec:efort_ellipsoid} - right), defined by:
\begin{equation} \label{ellipsoidforce}
\mathbf{f}^T (\mathbf{J}\mathbf{J}^T) \mathbf{f} \leq 1
\end{equation}

\subsubsection{Manipulability-based heuristics.}
From these definitions, we can derive three useful heuristics, that all account for the environment and the task being performed.
The first one, EFORT, was introduced by \cite{Tonneau2014}; the other two are new minor contributions, derived from these previous works.

@@

$h_{EFORT}$ and $h_{vel}$ will favor contacts that allow large efforts or fast modifications in the velocity.
$EFORT$ in particular is useful for tasks such as standing up, pushing / pulling.
In other less demanding cases, manipulability can also be considered to avoid singularities.
To do so, we can consider the manipulability measure $w$, also given by \citeauthor{Yoshikawa1984}:

\begin{equation*} \label{ellipsoid}
w(\mathbf{q}) = \sqrt{det(\mathbf{J}\mathbf{J}^T)}
\end{equation*}
$w$ measures the ``distance'' between a given configuration and a singular configuration. When $w$ is equal to 0, the configuration is in a singular state;
the greater $w$ is, the further away the configuration is from singularity.
We then derive:
\begin{equation}
h_{w}(\mathbf{q}, \mathbf{m}) = w(\mathbf{q})  (\mu_0 \mathbf{n}^T \mathbf{m})
\end{equation}

$h_{EFORT}$, $h_{vel}$ and $h_{w}$ define three kinematic heuristics, fast to compute (indeed, thanks to our sampling-based approach, the Jacobian and inverse products
can be precomputed offline), that allow the planner to select the best
candidates with respect to the local motion task along the guide path.
This results in faster computation times, and a reduced number of overall contact switches.
